Data Analysis of the Relationship Between Date, Air Temperature, Whether It Rains or Not, Water Temperature, and Observations of Water Fowls at Beaches in Toronto

### 2024-09-26 paper/paper.qmd
Update

---
title: "Data Analysis of Toronto Beach Data"
author: "Yanzun Jiang"
thanks: "Code and data are available at: LINK."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

@@

library(tidyverse)
library(palmerpenguins)
```

@@

The remainder of this paper is structured as follows. @sec-data....

@@

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false

ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = "none") +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

# analysis_data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))
# 
# analysis_data |> 
#   ggplot(aes(x = width, y = length)) +
#   geom_point(alpha = 0.8) +
#   theme_minimal() +
#   labs(x = "Wing width (mm)",
#        y = "Wing length (mm)")
```

Talk way more about it.

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

Our results are summarized in @tbl-modelresults.

@@

# Discussion

@@

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

>>>>

---
title: "Data Analysis of the Relationship Between Date, Air Temperature, Whether It Rains or Not, Water Temperature, and Observations of Water Fowls at Beaches in Toronto"
author: "Yanzun Jiang"
date: today
date-format: long
thanks: "Code and data supporting this analysis is available at: https://github.com/Stary54264/Toronto-Beach-Data"
abstract: "This paper conducted a basic data analysis on data of beaches in Toronto. The variables mentioned in the tile are highly correlated. The analysis can help us to know better about beaches in Toronto. Further, it helps meteorologists to monitor climate changes."
format: pdf
number-sections: true
bibliography: references.bib
---

@@

# load packages and read data
library(tidyverse)
library(knitr)
```

@@

The remainder of this paper is structured as follows.

@sec-data

&nbsp; @sec-data-1

&nbsp; @sec-data-2

&nbsp; &nbsp; @sec-data-2-1

&nbsp; &nbsp; @sec-data-2-2

&nbsp; &nbsp; @sec-data-2-3

&nbsp; &nbsp; @sec-data-2-4

@sec-results

@sec-discussion

@@

data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))
```

## Preview of data {#sec-data-1}

Our data (@tbl-data) is about beaches in Toronto, from @opendatatoronto.

```{r}
#| label: tbl-data
#| echo: false
#| warning: false
#| message: false

kable(head(data), format = "markdown")
```

This dataset records the date (which was separated into year, month, and day), the air temperature, whether it rains or not, the water temperature, and the observations of water fowls. We will analyze the relationship between these variables in the following part.

## Visualization of the Distribution of Each Observed Variable {#sec-data-2}

This part shows the distributions of each observed variables using histograms.

### Air Temperature {#sec-data-2-1}

This histogram (@fig-air-temp) shows the distribution of air temperatures of beaches in Toronto.

```{r}
#| label: fig-air-temp
#| fig-cap: Histogram of Air Temperature
#| echo: false
#| warning: false
#| message: false

data |>
  ggplot(aes(x = air_temp)) +
  geom_histogram(bins = 30, col = "darkblue", fill = "lightblue") +
  theme_minimal() +
  labs(x = "Air Temperature", y = "# of Occurences")
```

We can clearly see the data is unimodal and is not skewed.

### Rain {#sec-data-2-2}

This bar graph (@fig-rain) shows the distribution of whether it rains of not at beaches in Toronto.

```{r}
#| label: fig-rain
#| fig-cap: Bar Graph of Whether It Rains or Not
#| echo: false
#| warning: false
#| message: false

data |>
  ggplot(aes(x = rain)) +
  geom_bar(col = "darkblue", fill = "lightblue") +
  theme_minimal() +
  labs(x = "Whether It Rains or Not", y = "# of Occurences")
```

We can clearly see that there are less raining days.

### Water Temperature {#sec-data-2-3}

This histogram (@fig-water-temp) shows the distribution of water temperatures of beaches in Toronto.

```{r}
#| label: fig-water-temp
#| fig-cap: Histogram of Water Temperature
#| echo: false
#| warning: false
#| message: false

data |>
  ggplot(aes(x = water_temp)) +
  geom_histogram(bins = 30, col = "darkblue", fill = "lightblue") +
  theme_minimal() +
  labs(x = "Water Temperature", y = "# of Occurences")
```

We can clearly see the data is unimodal, which is the same as air temperatures, but it is left-skewed.

### Observations of Water Fowls {#sec-data-2-4}

This histogram (@fig-water-fowl) shows the distribution of observations of water fowls of beaches in Toronto.

```{r}
#| label: fig-water-fowl
#| fig-cap: Histogram of Observations of Water Fowls
#| echo: false
#| warning: false
#| message: false

data |>
  ggplot(aes(x = water_fowl)) +
  geom_histogram(bins = 30, col = "darkblue", fill = "lightblue") +
  theme_minimal() +
  labs(x = "Observations of Water Fowls", y = "# of Occurences")
```

We can clearly see the data is unimodal but highly right-skewed.

# Results {#sec-results}

Our results are summarized in.

@@

# Discussion {#sec-discussion}

@@
